---
title: "Consolidating Genre Variable"
author: "team 6"
date: "3/8/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This rmd contains the process used to clean the genre varaible and the anova used for the analysis

```{r packages, echo=FALSE}
library(broom)
library(knitr)
library(ggplot2)
```


```{r}
spotify <- read.csv("songs.csv")

genre <- as.factor(spotify$genre)

head(genre)

table(genre)
```


## Decreasing number of levels in genre 

Approach: Decrease the number of genres by categorizing specific genres into more broad ones 

Reasoning: Having less levels might make the genre variable more helpful

The table below has the most common genres:

```{r}
table(genre)[table(genre) > 50]
```

I grouped genre together based on key words in the genre names. I placed pop and hip hop at the bottom, as these two key words were in most of the genre levels.

hiphop: hip hop, R&B, 
electronic: Dance, electronic
rock: metal, rock 
country: country
latin: latin
pop: pop
other: reamining songs

```{r}
genre_new <- genre

rock_genre <- which(grepl(".*rock.*|.*metal.*", genre))

elec_genre <- which(grepl(".*Dance.*|.*Electronic.*", genre) & !grepl(".*rock.*|.*metal.*", genre))

hiphop_genre <- which( grepl(".*R\\&B.*|.*hip hop.*", genre) & !grepl(".*rock.*|.*metal.*|.*Dance.*|.*Electronic.*", genre) )


latin_genre <- which(grepl(".*latin.*", genre) & !grepl(".*rock.*|.*metal.*|.*Dance.*|.*Electronic.*|.*R\\&B.*|.*hip hop.*", genre))

pop_genre <- which(grepl(".*pop.*", genre) & !grepl(".*rock.*|.*metal.*|.*Dance.*|.*Electronic.*|.*R\\&B.*|.*hip hop.*|.*latin.*", genre))

other_genre <- which(!grepl(".*rock.*|.*metal.*|.*Dance.*|.*Electronic.*|.*R\\&B.*|.*hip hop.*|.*latin.*|.*pop.*", genre))

length(rock_genre) + length(elec_genre) + length(hiphop_genre) + length(latin_genre) + length(pop_genre) + length(other_genre)
```

Now we create the new column with less levels

```{r}

levels(genre_new) <- c(levels(genre_new), "other", "electric")

genre_new[rock_genre] <- "rock"
genre_new[elec_genre] <- "electric"
genre_new[hiphop_genre] <- "hip hop"
genre_new[latin_genre] <- "latin"
genre_new[pop_genre] <- "pop"
genre_new[other_genre] <- "other"

genre_new <- droplevels(genre_new)
table(genre_new)

# write.csv(data.frame(genre_new), file = "genre_new", row.names = F)

spotify$NewGenre <- genre_new
```

Here I run an Anova to find whether the genre of a song affects the danceability of the song

```{r}
dance_anova <- aov(danceability ~ NewGenre, data = spotify)
summary(dance_anova)

tapply(spotify$danceability, spotify$NewGenre, mean)

ggplot(spotify,aes(NewGenre,danceability, fill = NewGenre))+geom_boxplot()+ labs(title = "Danceability by Genre",
       x = "Genre",
       y = "Danceability")

```

|term      |   df|     sumsq|    meansq| statistic|p.value        |
|:---------|----:|---------:|---------:|---------:|:--------------|
|NewGenre  |    5|  5.106422| 1.0212844|  59.35865|< 2.220446e-16 |
|Residuals | 1994| 34.307400| 0.0172053|        NA|NA             |